A novel catalyitic model for estimating age-specific measles sero-prevalence uisng case-incidence data: a case study from the Democratic Republic of the Congo

### 2018-06-22 Paper/paper.tex
added title and authors
added links to missing references

\title{Catchy Title Goes Here}
\author{Laura Sampson}
\affiliation{Center for Infectious Disease Dynamics, Pennsylvania State University, State College, PA, 16801}
\author{Matthew Ferrari}
\affiliation{Center for Infectious Disease Dynamics, Pennsylvania State University, State College, PA, 16801}

@@

\section{Introduction \label{sec:Intro}}
Dramatic improvements in the coverage of vaccination with measles-containing vaccines have lead to significant reductions in the incidence of measles disease and the associated childhood mortality ~\cite{perry_global_2014}. Despite these improvements at the national scale, however, significant heterogeneity in both vaccination coverage and disease incidence remains at the sub-national level~\cite{prada_demographics_2017,takahashi_geography_2017,metcalf_transport_2015}.
Given the high transmission rate of measles, the proportion of the population that must be immunized to achieve and maintain elimination of measles is likely to be greater than 90\% in most populations [REF - Goodson?] (the exact level will be modulated by locally specific conditions such as population density and contact rates).
%"The WHO recommends that programs use good quality data to monitor population immunity by identifying and responding to large numbers of susceptibles" from Winter et al reference to WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28
%"Because the risk of measles outbreaks is determined by the rate of accumulation of susceptible people in the population, programmes should use available good quality data on population immunity (i.e. vacci- nation coverage, surveillance, serological studies) to monitor the accumulation of susceptible people and conduct follow-up campaigns before the number of pre-school children susceptible to measles approaches the equivalent of one birth cohort, in order to prevent an outbreak of measles. This approach has been found to be programmatically useful in preventing large outbreaks. However, for large countries and countries which are close to measles elimination, a more exten- sive assessment of the accumulation of susceptible persons should be carried out at the subnational level." WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28 -- http://apps.who.int/iris/bitstream/handle/10665/255149/WER9217.pdf?sequence=1

The size and age-distribution of the population susceptible to measles (hereafter we refer to this as the "susceptible persons") depends on the competing processes of susceptible recruitment through birth and susceptible loss through either natural infection or vaccination. 
Quantifying the size and age distribution of susceptible persons to measles is a critical tool in evaluating outbreak risk, the performance of vaccination programs, and developing vaccine-based interventions. The WHO recommends that programs monitor the accumulation of susceptible persons using "good quality data" for all countries, and that these should be made at the sub-national level for large countries and those close to measles elimination [REF -- Wkly Epid Rec].

@@

In the absence of a sero-survey, the number and age-distribution of susceptible persons can be estimated though demographic models that account for inputs through births and immunization via both vaccination and natural infection~\cite{METCALF2017S14,TRENTINI20171089,Winter2018Sero}. In this approach, the contribution of natural immunity is modeled as the result of a dynamic transmission model [REF] which requires explicit assumptions about the age-specific force of infection via a Who Acquires Infection From Whom (WAIFW) matrix that describes the age-specific rate of infectious contacts from each age class to each of the others.  Direct estimation of this WAIFW matrix is challenging from disease incidence or seroprevalence data (though see Whittaker and Farrington etc.), and thus it is common to make simplifying assumptions that either the WAIFW matrix is constant for all ages (i.e. a uniformly well-mixed system) ,or that the WAIFW matrix is a scalar function of some other measurable interaction process, such as contacts measured via diary studies~\ref{Polymod}.

>>>>

\title{A novel catalyitic model for estimating age-specific measles sero-prevalence uisng case-incidence data: a case study from the Democratic Republic of the Congo}
\author{Spencer Carran}
\affiliation{Center for Infectious Disease Dynamics, Pennsylvania State University, State College, PA, 16801}
\author{Steve Ahuka}
\affiliation{INRB, Kinshasa, DRC}
\author{Simon Bezirganian}
\affiliation{Center for Infectious Disease Dynamics, Pennsylvania State University, State College, PA, 16801}
\author{Matthew Ferrari}
\affiliation{Center for Infectious Disease Dynamics, Pennsylvania State University, State College, PA, 16801}
\author{Laura Sampson}
\affiliation{Center for Infectious Disease Dynamics, Pennsylvania State University, State College, PA, 16801}

@@

\section{Introduction \label{sec:Intro}}
Dramatic improvements in the coverage of vaccination with measles-containing vaccines have lead to significant reductions in the incidence of measles disease and the associated childhood mortality ~\cite{perry_global_2014}. Despite these improvements at the national scale, however, significant heterogeneity in both vaccination coverage and disease incidence remains at the sub-national level~\cite{prada_demographics_2017,takahashi_geography_2017,metcalf_transport_2015}.
Given the high transmission rate of measles, the proportion of the population that must be immunized to achieve and maintain elimination of measles is likely to be greater than 90\% in most populations [REF - https://www.ncbi.nlm.nih.gov/pubmed/21666163] (the exact level will be modulated by locally specific conditions such as population density and contact rates).
%"The WHO recommends that programs use good quality data to monitor population immunity by identifying and responding to large numbers of susceptibles" from Winter et al reference to WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28
%"Because the risk of measles outbreaks is determined by the rate of accumulation of susceptible people in the population, programmes should use available good quality data on population immunity (i.e. vacci- nation coverage, surveillance, serological studies) to monitor the accumulation of susceptible people and conduct follow-up campaigns before the number of pre-school children susceptible to measles approaches the equivalent of one birth cohort, in order to prevent an outbreak of measles. This approach has been found to be programmatically useful in preventing large outbreaks. However, for large countries and countries which are close to measles elimination, a more exten- sive assessment of the accumulation of susceptible persons should be carried out at the subnational level." WHO Measles vaccines: WHO position paper Wkly Epidemiol Rec 2017;92:205-28 -- http://apps.who.int/iris/bitstream/handle/10665/255149/WER9217.pdf?sequence=1

The size and age-distribution of the population susceptible to measles (hereafter we refer to this as the "susceptible persons") depends on the competing processes of susceptible recruitment through birth and susceptible loss through either natural infection or vaccination. 
Quantifying the size and age distribution of susceptible persons to measles is a critical tool in evaluating outbreak risk, the performance of vaccination programs, and developing vaccine-based interventions. The WHO recommends that programs monitor the accumulation of susceptible persons using "good quality data" for all countries, and that these should be made at the sub-national level for large countries and those close to measles elimination [REF -- https://www.ncbi.nlm.nih.gov/pubmed/28459148].

@@

In the absence of a sero-survey, the number and age-distribution of susceptible persons can be estimated though demographic models that account for inputs through births and immunization via both vaccination and natural infection~\cite{METCALF2017S14,TRENTINI20171089,Winter2018Sero}. In this approach, the contribution of natural immunity is modeled as the result of a dynamic transmission model [REF -- https://www.ncbi.nlm.nih.gov/pubmed/22178687] which requires explicit assumptions about the age-specific force of infection via a Who Acquires Infection From Whom (WAIFW) matrix that describes the age-specific rate of infectious contacts from each age class to each of the others.  Direct estimation of this WAIFW matrix is challenging from disease incidence or seroprevalence data (though see Whittaker and Farrington etc.), and thus it is common to make simplifying assumptions that either the WAIFW matrix is constant for all ages (i.e. a uniformly well-mixed system) ,or that the WAIFW matrix is a scalar function of some other measurable interaction process, such as contacts measured via diary studies~\ref{Polymod}.
